---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
  # pdf_document: default
  # fig_caption: yes
---

# truth
```{r}

this_truth_definition <- "found_in_3_of_4_biggest_subsets"

```


```{r setup, include = FALSE}

#knitr::opts_knit$set(root.dir = "~/Documents/Dropbox/ucsc/projects/gitCode/calculations for umend_qc_publication/")

library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(forcats)

```

## Load data

```{r}

outlier_results_raw <- read_tsv("../data/outliers.txt") %>% mutate(is_outlier_in_subset = TRUE)

this_bin_width = 4

evenly_selected_subsamples <- read_tsv(paste0("../data/evenly_selected_subsamples_with_binwidth_4.tsv"))

parent_metadata <- read_tsv("../data/parent_sample_data_source.tsv")

literature_outliers <- read_tsv("../ancillary/genes_with_journal_article_hit_counts.tsv") %>%
  left_join(parent_metadata %>% select(Eval_ID, Disease), by=c("disease"="Disease"))

```

## Definitions

```{r}

total_genes_measured <- 58582*0.05

resultsCodes <- tibble(gene_is_true_positive_in_parent_sample = c(T, F, T, F),
                       is_outlier_in_subset = c(T,T,F,F),
                       hit_type = c("True positive", "False positive", "False negative", "True negative"))

source_palette <- brewer.pal(12, "Paired")

cb_dark_red <- source_palette[6]
cb_light_red <- source_palette[5]

cb_dark_blue <- source_palette[2]
cb_light_blue <- source_palette[1]

plot_colors <- c("False positive" = cb_light_red,
                "True negative" = cb_dark_red,
                "True positive" = cb_dark_blue,
                "False negative" = cb_light_blue)

outlier_line_colors <- c("sensitivity" = as.character(plot_colors["True positive"]),
                          "specificity" = as.character(plot_colors["True negative"]))


```

## Identify putative true outliers

```{r}


annotated_outlier_results <- lapply(unique(outlier_results_raw$parent_sample)[2:5], function(this_parent_sample) {
	# this_parent_sample <- unique(outlier_results_raw$parent_sample)[2]
	outliers_found_in_selected_subsamples <- subset(outlier_results_raw,
	                                                parent_sample == this_parent_sample &
	                                                  sample_id %in% evenly_selected_subsamples$sample_id)
	subsamples <- unique(outliers_found_in_selected_subsamples$sample_id)
	genes_that_are_outliers_in_any_subset <- unique(outliers_found_in_selected_subsamples$gene)
	# add data to show a negative result for genes not found in the outlier list
	empty_results <- tibble(parent_sample = this_parent_sample,
	                        sample_id = rep(subsamples,
	                                        each = length(genes_that_are_outliers_in_any_subset)),
	                        gene = rep(genes_that_are_outliers_in_any_subset,
	                                   length(subsamples)))
	outlier_results <- left_join(empty_results,
	                             outliers_found_in_selected_subsamples,
	                             by = c("parent_sample", "sample_id", "gene")) %>%
	  mutate(target_MEND_count = as.numeric(gsub("^.*_est", "",
	                                              gsub("M_.MEND_seqtk_seed_.*", "", sample_id))),
	         is_outlier_in_subset = ifelse(is.na(is_outlier_in_subset), FALSE, TRUE))
	
	# define true positives

	if (this_truth_definition == "found_in_3_of_4_biggest_subsets") {
  	biggest_4_subsets <- sort(unique(outlier_results$target_MEND_count), decreasing = TRUE)[1:4]
  	minimum_hits_for_true_outlier <- 3
    
    outlier_frequency <- outlier_results %>% 
      dplyr::filter(is_outlier_in_subset,
                    target_MEND_count %in% biggest_4_subsets) %>% 
      group_by(gene) %>% 
      summarize(n_subsamples_found = length(unique(sample_id)))
    
    true_outliers_in_parent_sample <- outlier_frequency %>% 
      dplyr::filter(n_subsamples_found >= minimum_hits_for_true_outlier) %>% 
      pull(gene)
  }
		
	# except when True = present@50, exclude candidate outliers that are never positive in this data set
  if (this_truth_definition != "found_in_the_deepest_subset") {
	  outlier_results <- outlier_results %>% 
  	  group_by(gene) %>%
  	  dplyr::filter(sum(is_outlier_in_subset) > 0) 
  }
	                          
	 this_annotated_outlier_results <- outlier_results %>% group_by(gene) %>% 
		mutate(up_outlier_occurances = sum(as.numeric(is_outlier_in_subset)),
		       first_appears = min(target_MEND_count[is_outlier_in_subset]),
		       gene_is_true_positive_in_parent_sample = gene %in% true_outliers_in_parent_sample) %>%
	   left_join(resultsCodes, by = c("is_outlier_in_subset", "gene_is_true_positive_in_parent_sample"))
	 return(this_annotated_outlier_results)
}) %>% bind_rows
```

# Load t-test results
```{r}

cluster_results_matrix = matrix(
  c("medulloblastoma", "Medulloblastoma_Group3or4_v10_ttest.tsv",
  "embryonal rhabdomyosarcoma", "EmbryonalRhabdomyosarcoma_v10_ttest.tsv",
  "glioma (astrocytoma)", "Glioma_Astrocytoma_v10_ttest.tsv"),
            ncol=2, byrow=TRUE)
colnames(cluster_results_matrix) = c("disease", "file")
cluster_results_files <- as_tibble(cluster_results_matrix)

  

library(purrr)
cluster_results <- cluster_results_files %>% 
  mutate(file_contents = map(file, read_tsv)) %>%
  unnest %>%
  left_join(parent_metadata %>% select("Disease", "Eval_ID"), by=c("disease"="Disease"))



default_alpha <- 0.05 # p-value threshold
n_diseases <- nrow(cluster_results_files)

paste(sum(cluster_results$Pvalue<default_alpha), "genes in", n_diseases, "diseases have a p value below", default_alpha)

# apply bonferroni correction per disease
revised_alpha_threshold = default_alpha/nrow(cluster_results)

cluster_results_above_bonf_p <- cluster_results %>%
  filter(Pvalue<revised_alpha_threshold)

paste(nrow(cluster_results_above_bonf_p), "genes in", n_diseases, "diseases have a p value below the bonf corrected p value of", formatC(revised_alpha_threshold, format = "e", digits = 2))


# take the median number of outliers found for a sample
median_outlier_count <- 200 # estimate

cluster_supported_outliers <- cluster_results_above_bonf_p %>%
  group_by(disease) %>%
  top_n(median_outlier_count, tstats) %>%
  ungroup %>%
  arrange(desc(tstats)) %>% 
  mutate(has_cluster_support = TRUE) %>%
  select(Eval_ID, Gene, has_cluster_support)


subset(cluster_supported_outliers, Gene == "OTX2")

```

# Analyze medullo t-test results
```{r}
# medullo_scores <- these_scores
# subset(medullo_scores, Gene == "OTX2")
# 
# # survey p values
# hist(medullo_scores$Pvalue)
# nrow(medullo_scores) # 27,222
# 
# # how many pass significance 
# default_alpha <- 0.05 # p-value threshold
# sum(medullo_scores$Pvalue<default_alpha)  #12,261 -- unreasonably big
# 
# # assess effect of bonferroni correction
# revised_alpha_threshold = default_alpha/nrow(medullo_scores)
# sum(medullo_scores$Pvalue<revised_alpha_threshold) #8054 -- still unreasonably big
# 
# 
# # take the median number of outliers found for a sample
# median_outlier_count <- 200 # estimate
# 
# medullo_scores %>%
#   top_n(median_outlier_count, tstats) %>%
#   arrange(desc(tstats))

```

# Filter for hits supported by t-test results of clusters
```{r}

annotated_outlier_results_intermediate <- annotated_outlier_results %>%
  left_join(evenly_selected_subsamples %>% select(-target_MEND_count), 
            by = "sample_id") %>%
  left_join(cluster_supported_outliers, by=c("parent_sample" = "Eval_ID", "gene"="Gene")) %>%
  mutate(has_cluster_support = ifelse (is.na(has_cluster_support), FALSE, TRUE)) %>%
  filter(has_cluster_support)


# report overlap: 
  
annotated_outlier_results_intermediate %>%
  select(parent_sample, gene) %>%
  distinct %>%
  group_by(parent_sample) %>%
  summarize(n_genes=n())

# annotated_outlier_results$bin_median <- evenly_selected_subsamples$median_of_depth_bin[match(
#   annotated_outlier_results$sample_id, evenly_selected_subsamples$sample_id)]

ordered_plot_input <- annotated_outlier_results_intermediate %>% 
  arrange(gene_is_true_positive_in_parent_sample, up_outlier_occurances, rev(first_appears)) %>%
  ungroup %>%
  mutate(unique_parent_gene_id = paste(parent_sample, gene, sep = "_"),
         ordered_gene = factor(unique_parent_gene_id, levels = rev(unique(unique_parent_gene_id)))) %>%
  group_by(parent_sample) %>%
  mutate(parent_sample_label = paste0(parent_sample, " (n=", length(unique(unique_parent_gene_id)), ")"))

```


# Make minimal plot 
```{r  fig.height=12, fig.width=10}


outlier_results_x_breaks=seq(
        0, 
        max(ordered_plot_input$median_of_depth_bin),
        by=this_bin_width
      )
outlier_results_x_labels=gsub("^0$", 1, outlier_results_x_breaks)


plot_title="outliers - each row is a gene"  
# pIA<-
  
  ggplot(subset(ordered_plot_input), aes(x = median_of_depth_bin, y = ordered_gene, fill = hit_type)) + 
  geom_tile(stat = "identity") +
  scale_fill_manual("Call", values= plot_colors) +
#  ggtitle(plot_title) + theme(axis.text.y  = element_blank())+
#  scale_x_continuous("MEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) +
  xlab("MEND reads (million)") + 
  ylab("Genes") + 
  facet_grid(parent_sample ~ . , scales="free_y", space="free") +
  theme(
   # axis.text.y  = element_blank(),
    axis.ticks.y  = element_blank()
  )  + 
  theme(legend.position="none")
  
  
# pIA
# ggsave(paste0("figures_and_tables/", plot_title,".", Sys.Date(),  ".png"), plot=pIA, width=3, height=3)
```
## Calculate fractional accuracy

```{r}

frac_accuracy <- ordered_plot_input %>%
  group_by(parent_sample, sample_id, median_of_depth_bin, parent_sample_label, publication_id) %>%
	summarize(fp_frac = sum(hit_type == "False positive")/n(),
	          tn_frac = sum(hit_type == "True negative")/n(),
	          fn_frac = sum(hit_type == "False negative")/n(),
	          tp_frac = sum(hit_type == "True positive")/n())
		
conv_table1 <- tibble(fraction_type = c("fp_frac", "tn_frac", "fn_frac", "tp_frac"),
                      hit_type = c("False positive", "True negative", "False negative", "True positive"))
	
g_frac_accuracy <- gather(frac_accuracy,
                          fraction_type,
                          fraction_amount,
                          -sample_id,
                          -parent_sample,
                          -median_of_depth_bin,
                          -parent_sample_label,
                          -publication_id) %>%
  left_join(conv_table1) %>%
	mutate(hit_type = factor(hit_type, levels = c("True negative",
	                                              "False positive",
	                                              "False negative",
	                                              "True positive")))

head(g_frac_accuracy)



```



## Calculate sensitivity and specificity

```{r, fig.height = 10, fig.width = 9}

colsToGather = c("median_of_depth_bin", "sensitivity", "specificity", "ppv", "npv")

accuracy_stats <- NULL
for (this_parent_sample in unique(ordered_plot_input$parent_sample)) {
  # this_parent_sample = "TH_Eval_018"
  this_plot_input <- subset(ordered_plot_input, parent_sample == this_parent_sample)
  count_of_positives_in_full_set <- length(unique(subset(this_plot_input,
                                                         gene_is_true_positive_in_parent_sample)$gene))
  count_of_negatives_in_full_set <- total_genes_measured-count_of_positives_in_full_set
  genes_ever_called_as_outlier <- unique(this_plot_input$gene)
  count_of_genes_not_in_outlier_list <- total_genes_measured-length(genes_ever_called_as_outlier)
    
  # calculate statistics to compare across samples
  accuracy_counts<-this_plot_input %>% 
    group_by(median_of_depth_bin) %>% 
    summarize(all_positives_in_subset = sum(is_outlier_in_subset), 
              all_negatives_in_subset = sum(!is_outlier_in_subset) + count_of_genes_not_in_outlier_list, 
              tp_count_in_subset = sum(hit_type == "True positive"), 
              fp_count_in_subset = sum(hit_type == "False positive"),
              partial_tn_count_in_subset = sum(hit_type == "True negative"), 
              total_tn_count_in_subset = partial_tn_count_in_subset + count_of_genes_not_in_outlier_list,
              fn_count_in_subset = sum(hit_type == "False negative"))
    
  accuracy_stats_this_parent <- accuracy_counts %>% 
    mutate(sensitivity = tp_count_in_subset / count_of_positives_in_full_set,
           specificity = total_tn_count_in_subset / count_of_negatives_in_full_set, 
           ppv = tp_count_in_subset / all_positives_in_subset,
           npv = total_tn_count_in_subset / all_negatives_in_subset)
    
  fullStatNames = data.frame(short = c("sensitivity", "specificity", "ppv", "npv"),
                             long = c("sensitivity-detect when present",
                                      "specificity-exclude when absent",
                                      "ppv-positive predictive value",
                                      "npv-negative predictive value"))
  
  g_accuracy_stats_this_parent = gather(accuracy_stats_this_parent[,
                                                                   colsToGather],
                                        measurement,
                                        value,
                                        -median_of_depth_bin) %>%
    mutate(parent_sample=this_parent_sample, 
           publication_id = unique(this_plot_input$publication_id))
  
  accuracy_stats <- bind_rows(accuracy_stats, g_accuracy_stats_this_parent)
}



accuracy_stats_sens_spec<- accuracy_stats %>% 
  dplyr::filter(measurement %in% c("sensitivity", "specificity")) %>%
  mutate(median_of_depth_bin = ifelse(median_of_depth_bin==0, 1, median_of_depth_bin)) %>%
  group_by(measurement, median_of_depth_bin) %>%
  mutate(
    min_value = min(value),
    # publication_id=parent_publication_ids$pub_ID[match(parent_sample, parent_publication_ids$Eval_ID)],
    is_median_point=c("not the median", "median value")[1+(value==median(value))]
         )



```


# plot specifications 
```{r}


outlier_results_x_breaks=seq(
        0, 
        max(ordered_plot_input$median_of_depth_bin),
        by=this_bin_width
      )
outlier_results_x_labels=gsub("^0$", 1, outlier_results_x_breaks)

```


### outliers - each row is a gene
```{r fig.height=12, fig.width=10}


plot_title="outliers - each row is a gene"  
# pIA<-
  
  ggplot(subset(ordered_plot_input), aes(x = median_of_depth_bin, y = ordered_gene, fill = hit_type)) + 
  geom_tile(stat = "identity") +
  scale_fill_manual("Call", values= plot_colors) +
#  ggtitle(plot_title) + theme(axis.text.y  = element_blank())+
#  scale_x_continuous("MEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) +
  xlab("MEND reads (million)") + 
  ylab("Genes") + 
  facet_grid(parent_sample ~ . , scales="free_y", space="free") +
  theme(
   # axis.text.y  = element_blank(),
    axis.ticks.y  = element_blank()
  )  + 
  theme(legend.position="none")
  
  
# pIA
# ggsave(paste0("figures_and_tables/", plot_title,".", Sys.Date(),  ".png"), plot=pIA, width=3, height=3)
```


## disease and sample
```{r}
parent_metadata %>% select(Eval_ID, Disease) %>% kable


```


### Figure part C - sensitivity and specificity

```{r}
theme_set(theme_gray())

plot_title <- "outliers - sensitivity and specificity"  


pOSS <- ggplot(accuracy_stats_sens_spec, aes(y = value, color = measurement, x = median_of_depth_bin)) +
  geom_point() +
  geom_smooth(se = TRUE) + 
  xlab("MEND reads (million)") + 
  ylab("Fraction") +
  facet_grid(publication_id~ . ) +
  scale_color_manual("Measurement", values = outlier_line_colors) +
 scale_y_continuous(breaks=c(0.8, 0.9, 1))  + 
  theme(legend.position="none",
        strip.text.y = element_text(angle = 0))

pOSS

# ggsave(paste0("figures_and_tables/", plot_title,".", Sys.Date(),  ".png"), plot=pOSS, width=3, height=6)

```


	
# Table of accuracy stats

```{r}
# 
# accuracy_stats_combo_for_pub <- accuracy_stats_sens_spec %>% 
#   select(depth_bin_min_value=min_value, everything()) %>%
#   select(median_of_depth_bin,measurement, depth_bin_min_value) %>%
#   distinct() %>%
#   spread(measurement, depth_bin_min_value) 
# 
# 
# accuracy_stats_combo_for_pub %>% 
#   kable( digits=4)
# 
# write_tsv(accuracy_stats_combo_for_pub, "../data/outlier_minimum_accuracy_stats_combo_for_pub.tsv")

View(accuracy_stats_sens_spec)

accuracy_stats_sens_spec[, 1:4] %>%
  spread(measurement, value) %>%
  arrange(parent_sample, median_of_depth_bin)

median_sens_spec_values<-accuracy_stats_sens_spec %>%
  dplyr::filter(!is_median_point=="not the median") %>%
  select(median_value=value, everything()) %>%
  select(median_of_depth_bin,measurement, median_value) %>%
  distinct() %>%
  spread(measurement, median_value) 

median_sens_spec_values %>% 
  kable( digits=4)

# write_tsv(median_sens_spec_values, "../data/outlier_median_accuracy_stats_combo_for_pub.tsv")

```