---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  toc: true
  # pdf_document: default
  # fig_caption: yes
---

In this analysis, the true expression of a gene is considered to be the expression level found_in_the_deepest_subset made from that parent. An accurate measurement is defined as one that is within 25% percent of the true expression values. 

# Set submittable print options

```{r}

f_final_draft <- TRUE
notation_text_alpha <- ifelse (f_final_draft, 0, 1 ) 

```


```{r setup, include = FALSE}

library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(forcats)
library(cowplot)

```

## Load data

```{r}

gene_abundances_for_bin_detail <- read_tsv("../data/gene_abundances_for_bin_detail.tsv")

highlighted_gene_abundances_raw <- read_tsv("../data/highlighted_gene_abundances.tsv")

read_counts_raw <- read_tsv("../data/read_counts.txt",
                            col_names = c("sample_id", "measurement", "read_count"))

```

## Groom data

```{r}

 
# add information about total reads at data points
highlight_expression <- left_join(highlighted_gene_abundances_raw,
                            read_counts_raw %>%
                              dplyr::filter(measurement == "total_sequences") %>%
                              select(-measurement) %>%
                              select(total_reads = read_count, everything()),
                            by = "sample_id") 

```
 
# Set grid line sizes

```{r}

grid_size_major <- 0.08
grid_size_minor <- 0.05

```

## Highlight expression of some cancer genes

```{r, fig.height = 5, fig.width = 9}

expression_label_data <- highlight_expression %>%
  dplyr::filter(
    parent_id == "TH_Eval_014",
    target_MEND_count == max(target_MEND_count)) %>% 
  mutate(vjust_val = ifelse(gene %in% c("JAK2", "TP53"), 2, -0.5))

# max_x=20+10*ceiling(max(expression_label_data$max_depth)/1e7)
max_x <- 10 * ceiling(max(expression_label_data$total_reads) / 1E7)

p <- ggplot(highlight_expression) +
  geom_line(aes(x = total_reads / 1e6, y = expression, group = gene, color = gene)) +
  geom_point(aes(x = total_reads / 1e6, y = expression, group = gene, color = gene), size = 0.5)
  # shape= within_accuracy_threshold_of_deepest_val),  size=3, fill="white", alpha=.8 # line opts were , size=1.5, alpha=.7

p2 <- p +
  geom_text(data = expression_label_data,
            aes(x = max_x - 5,
                y = expression,
                label = gene,
                color = gene,
                vjust = vjust_val),
            size = 4, hjust = 1) #+
  #expand_limits(x = max(expression_label_data$max_depth) / 1e6+5)  

p3 <- p2 + 
  facet_wrap(~publication_id, nrow = 1) + 
  background_grid(major = "xy", 
                  minor = "x",
                  colour.major = "black", 
                  colour.minor = grey(0.75),
                  size.major = grid_size_major,
                  size.minor = grid_size_minor) + 
  theme(axis.line = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0)) + 
  scale_shape_manual(values = c(22, 16))  + 
  scale_color_brewer(guide = FALSE, palette = "Set1") + 
  expand_limits(y = 0, x = max_x) +
  ylab("Expression (log2(TPM+1))") + 
  xlab("Total reads (million)") +
  ggtitle("A")

p3

```

# Smaller

```{r}

ids_to_plot <- c("S1", "S2")
small_highlight_expression <- subset(highlight_expression, publication_id %in% ids_to_plot)

p <- ggplot(small_highlight_expression) +
  geom_line(aes(x = total_reads / 1e6,
                y = expression,
                group = gene,
                color = gene)) +
  geom_point(aes(x = total_reads / 1e6,
                 y = expression,
                 group = gene,
                 color = gene),
             size = 0.5)  # shape= within_accuracy_threshold_of_deepest_val),  size=3, fill="white", alpha=.8 # line opts were , size=1.5, alpha=.7

p2 <- p +
  geom_text(data = expression_label_data,
            aes(x = 1 + total_reads / 1e6,
                y = expression,
                label = gene,
                color = gene,
                vjust = vjust_val),
            size = 4,
            hjust = 1) #+
  #expand_limits(x = max(expression_label_data$max_depth) / 1e6 + 5)  

small_p3 <- p2 + 
  facet_wrap(~publication_id, nrow = 1) + 
  background_grid(major = "xy", 
                  minor = "x",
                  colour.major = "black", 
                  colour.minor = grey(0.75),
                  size.major = grid_size_major,
                  size.minor = grid_size_minor) + 
  theme(axis.line = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0)) + 
  scale_shape_manual(values = c(22, 16))  + 
  scale_color_brewer(guide = FALSE, palette = "Set1") + 
  expand_limits(y = 0, x = max_x) +
  ylab("Expression (log2(TPM+1))") + 
  xlab("Total reads (million)") +
  ggtitle("A")

small_p3

```

# Combine accuracy across expression bins

These values are not used in calculations here, just for reporting. They are used in "summarize_expression"
```{r}

half_size_of_expr_bin <- 0.02
gene_expression_levels <- c(1.5, 3.5, 5.5)


n_gene_sample_combos <- gene_abundances_for_bin_detail %>% 
  filter(target_MEND_count == max(target_MEND_count)) %>%
  group_by(parent_id, expression_bin) %>%
  summarize(n_genes_per_boxplot = n())

unique(n_gene_sample_combos$n_genes_per_boxplot)

```

#  Box plots across expression bins by total reads

note: 
each expression bin contains `r unique(n_gene_sample_combos$n_genes_per_boxplot)` genes per parent sample
Each expression bin contains genes that are expressed within +/- `r half_size_of_expr_bin` of `r gene_expression_levels`

# All five parent samples

```{r}

pretty_expression <-  left_join(gene_abundances_for_bin_detail,
                            read_counts_raw %>%
                              dplyr::filter(measurement == "total_sequences") %>%
                              select(-measurement) %>%
                              select(total_reads = read_count, everything()),
                            by = "sample_id") 



# dummy data to make the y axis as big as i want them
dummy <- bind_rows(c(median_expr = 1.5, expression = 5),
                   c(median_expr = 1.5, expression = 0),
                   c(median_expr = 3.5, expression = 6),
                   c(median_expr = 3.5, expression = 1),
                   c(median_expr = 5.5, expression = 3),
                   c(median_expr = 5.5, expression = 8)) %>%
  mutate(publication_id = "S1", # it doesn't matter which since they're all using the same y axis
         total_reads = 10 * 1e6)

pretty_expression$median_expr_f <- factor(as.character(pretty_expression$expression_bin),
                                          levels = rev(as.character(gene_expression_levels)))
dummy$median_expr_f <- factor(as.character(dummy$median_expr),
                              levels = rev(as.character(gene_expression_levels)))

hline_dummy <- as_tibble(expand.grid(median_expr = unique(dummy$median_expr),
                                     publication_id = unique(pretty_expression$publication_id))) %>%
  mutate(median_expr_f = factor(as.character(median_expr)))

plot_title = "Range of expression of genes with similar expression"
pBin <- ggplot(pretty_expression) + 
  geom_hline(data = hline_dummy, aes(yintercept = median_expr),
             size = 0.25, color = "#6699cc") +
  geom_boxplot(aes(x = total_reads / 1e6,
                   y = expression,
                   group = paste(expression_bin, median_of_depth_bin)), #color=label,
               outlier.size = 0.5, width = 3) + 
  xlab("Total reads (million)") + 
  ylab("Expression (log2(TPM+1))") + 
  theme(legend.position = "none") + 
  facet_grid(median_expr_f ~ publication_id, scales = "free_y") +
  geom_blank(data = dummy, aes(x = total_reads / 1e6, y = expression)) +
  scale_y_continuous(breaks = seq(0, 10, by = 2)) + 
  ggtitle("B")
  # ggtitle(plot_title)

pBin2 <- pBin +
  background_grid(major = "xy", 
                  minor = "x",
                  colour.major = "black", 
                  colour.minor = grey(0.75),
                  size.major = grid_size_major,
                  size.minor = grid_size_minor) +
  theme(axis.line = element_blank(),
        plot.title = element_text(hjust = 0)) +
  scale_x_continuous(breaks = seq(0, max_x, 25))
  # minor_breaks = seq(5, max_x, 5)) +
  # scale_x_continuous(breaks = seq(0, max_x, 25),
  # minor_breaks = seq(5, max_x, 5)) +

pBin2
# ggsave(paste0("figures_and_tables/", plot_title, ".png"), plot=pBin, width=7, height=7)

```

# Just 2 parent samples

```{r}

# dummy data to make the y axis as big as i want them
dummy <- bind_rows(c(expression_bin = 1.5, expression = 5),
                   c(expression_bin = 1.5, expression = 0),
                   c(expression_bin = 3.5, expression = 6),
                   c(expression_bin = 3.5, expression = 1),
                   c(expression_bin = 5.5, expression = 3),
                   c(expression_bin = 5.5, expression = 8)) %>%
  mutate(publication_id = "S1", # it doesn't matter which since they're all using the same y axis
         total_reads = 10 * 1e6)

hline_dummy <- as_tibble(expand.grid(expression_bin = unique(dummy$expression_bin),
                                     publication_id = unique(pretty_expression$publication_id))) %>%
  mutate(
    expression_bin_f = factor(as.character(expression_bin), as.character(gene_expression_levels))
  )

pretty_expression$expression_bin_f <- factor(as.character(pretty_expression$expression_bin),
                                             levels = as.character(gene_expression_levels))
dummy$expression_bin_f <- factor(as.character(dummy$expression_bin),
                              levels = as.character(gene_expression_levels))
hline_dummy$expression_bin_f <- factor(hline_dummy$expression_bin_f,
                                    levels = as.character(gene_expression_levels))

ids_to_plot <- c("S1", "S2")
small_pretty_expression <- subset(pretty_expression, publication_id %in% ids_to_plot)
small_hline_dummy <- subset(hline_dummy, publication_id %in% ids_to_plot)

plot_title <- "Range of expression of genes with similar expression"
pBin <- ggplot(small_pretty_expression) + 
  geom_hline(data = small_hline_dummy, aes(yintercept = expression_bin),
             size = 0.25, color = "#6699cc") + # color = "grey") +
  geom_boxplot(aes(x = total_reads / 1e6,
                   y = expression,
                   group = paste(expression_bin, median_of_depth_bin)), #color=label,
               outlier.size = 0.25, width = 3) + 
  # geom_hline(data=hline_dummy, aes(yintercept=expression_bin), size=0.5, linetype="dotted", color="blue") +
  xlab("Total reads (million)") + 
  ylab("Expression (log2(TPM+1))") + 
  theme(legend.position = "none") + 
  facet_grid(expression_bin_f ~ publication_id, scales = "free_y") +
  geom_blank(data = dummy, aes(x = total_reads / 1e6, y = expression)) +
  scale_y_continuous(breaks = seq(0, 10, by = 2)) + 
  ggtitle("B")
  # ggtitle(plot_title)

pSmallBin2 <- pBin +
  background_grid(major = "xy", 
                  minor = "x",
                  colour.major = "black", 
                  colour.minor = grey(0.75),
                  size.major = grid_size_major,
                  size.minor = grid_size_minor) +
  theme(axis.line = element_blank(),
        plot.title = element_text(hjust = 0)) +
  scale_x_continuous(breaks = seq(0, max_x, 25))

pSmallBin2

```

## Combine plots

```{r}

combined_figure_name <- "expression_figures"

library(grid)
library(gtable)
library(gridExtra)

grob1 <- ggplotGrob(p3)
grob2 <- ggplotGrob(pBin2)

this_time_stamp <- Sys.time()

grobCombo <- arrangeGrob(grobs = list(grob1, grob2), 
                         layout_matrix = rbind(c(1,1),
                                               c(1,1),
                                               c(2,2),
                                               c(2,2),
                                                c(2,2)),
                         bottom = textGrob(paste(combined_figure_name, this_time_stamp, "UTC"),
                                           x=0,
                                           just = "left",
                                           gp = gpar(fontsize = 8, alpha = notation_text_alpha)))
grid.newpage()
grid.draw(grobCombo)

ggsave(file = paste0("../figures_and_tables/",
                     combined_figure_name,
                     "_",
                     gsub(" ", "_", gsub(":", "-", this_time_stamp)),
                     "_",
                     "UTC",
                     ".png"),
       grobCombo,
       width = 6,
       height = 8)


ggsave(file = paste0("../figures_and_tables/",
                     combined_figure_name,
                     "_",
                     gsub(" ", "_", gsub(":", "-", this_time_stamp)),
                     "_",
                     "UTC",
                     ".tiff"),
       grobCombo,
       width = 6,
       height = 8)
```

## Combine small plots

```{r}

combined_figure_name <- "two_sample_expression_figure"

library(grid)
library(gtable)
library(gridExtra)

grob1 <- ggplotGrob(small_p3)
grob2 <- ggplotGrob(pSmallBin2)

this_time_stamp <- Sys.time()

grobCombo <- arrangeGrob(grobs = list(grob1, grob2), 
                         layout_matrix = rbind(c(1,2),
                                               c(1,2)),
                         bottom = textGrob(paste(combined_figure_name, this_time_stamp, "UTC"),
                                           x = 0,
                                           just = "left",
                                           gp = gpar(fontsize = 8, alpha = notation_text_alpha)))
grid.newpage()
grid.draw(grobCombo)

ggsave(file = paste0("../figures_and_tables/",
                     combined_figure_name,
                     "_",
                     gsub(" ", "_", gsub(":", "-", this_time_stamp)),
                     "_",
                     "UTC",
                     ".png"),
       grobCombo,
       width = 8,
       height = 4)

ggsave(file = paste0("../figures_and_tables/",
                     combined_figure_name,
                     "_",
                     gsub(" ", "_", gsub(":", "-", this_time_stamp)),
                     "_",
                     "UTC",
                     ".tiff"),
       grobCombo,
       width = 8,
       height = 4)

```

# Statements about analysis

format: With 1 million reads, the values of the five measured genes ranged from 65 to 577% of the values measured at the maximum depth. With 5 million reads, the range is 0 to 132%. WIth 10 million reads, 88 to 120%.

```{r}


max_depth=max(highlight_expression$target_MEND_count)

highlight_expression_for_statements <- 
  highlight_expression %>% 
  group_by(gene, parent_id) %>%
  mutate(
    expression_at_max_depth = expression[target_MEND_count==max_depth],
    percent_of_max_value = 100*expression/expression_at_max_depth)


# 1 million MEND reads is roughtly 1 million total reads
this_depth <- 1
highlight_expression %>%
  filter(target_MEND_count==this_depth) %>%
  pull(total_reads) %>%
  range

highlight_expression_for_statements %>%
  filter(target_MEND_count==this_depth) %>%
  arrange(percent_of_max_value) %>%
  select(-median_of_depth_bin, -sample_id) %>%
  pull(percent_of_max_value) %>%
  range %>%
  round


# 4 million MEND reads is roughtly 5 million total reads
this_depth <- 4
highlight_expression %>%
  filter(target_MEND_count==this_depth) %>%
  pull(total_reads) %>%
  range

highlight_expression_for_statements %>%
  filter(target_MEND_count==this_depth) %>%
  arrange(percent_of_max_value) %>%
  select(-median_of_depth_bin, -sample_id) %>%
  pull(percent_of_max_value) %>%
  range %>%
  round


# 8 million MEND reads is roughtly 10 million total reads
this_depth <- 8
highlight_expression %>%
  filter(target_MEND_count==this_depth) %>%
  pull(total_reads) %>%
  range

highlight_expression_for_statements %>%
  filter(target_MEND_count==this_depth) %>%
  arrange(percent_of_max_value) %>%
  select(-median_of_depth_bin, -sample_id) %>%
  pull(percent_of_max_value) %>%
  range %>%
  round


```
