---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_knit$set(root.dir = "~/Documents/Dropbox/ucsc/projects/gitCode/drafts_of_umend_qc_publication/")

library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(gridGraphics)
library(RColorBrewer)
library(viridis)
library(ggrepel)
library(cowplot)

```

# Settings

```{r}

n_samples_to_plot <- 50

library(RColorBrewer)

source_palette <- brewer.pal(12, "Paired")
cb_dark_green <- source_palette[4]
cb_light_green <- source_palette[3]



raw_names_for_read_types <-  c("not_uniquely_mapped", "duplicate_reads", "non_exonic_reads", "UMEND")
intermediate_names_for_read_types = c("Not uniquely mapped", "Duplicate reads", "Non exonic reads", "UMEND")
better_names_for_read_types <- c("Not uniquely mapped", "Duplicate", "Non exonic", "UMEND")


category_colors <- c(grey(0.5), cb_light_green, grey(0.7), cb_dark_green) 
names(category_colors) <- intermediate_names_for_read_types

```

# Load data

```{r}

allCountsRaw <- read_tsv(file.path( "../data/raw_read_survey_results_2019_03_28.txt"), 
                         col_names = c("sample_id", "count_type", "read_count")) %>%
  spread(count_type, read_count) 

parent_sample_data <- read_tsv("../data/parent_sample_data_source.tsv") %>% rename(pub_ID = publication_id)

```

# Groom data 

```{r}

complete_counts_raw <- na.omit(allCountsRaw) %>%
  filter(!grepl("TH06", sample_id)) # exclude placeholder QC values from TH06 (we don't have real QC values for these samples)




# consider excluding samples in which duplicates may have been removed
# dplyr::filter(!grepl("THR08", sample_id )) # i think duplicates may have been removed from these samples
# nrow (filter(allCountsRaw, grepl("THR08", sample_id )))


# effect of excluding NA, which are usually samples without four measurements
nrow(na.omit(allCountsRaw))
nrow((allCountsRaw))


```

## Simplify names

```{r}

read_counts <- complete_counts_raw %>%
  rename(UM = Uniquely_mapped_read_count,
         UMND = Uniquely_mapped_non_duplicate_read_count)

```

## Calculate fractions of read types

```{r}

orderByTotalReads <- read_counts %>% arrange(desc(total_sequences)) %>%	.$sample_id
read_counts_with_slices <- read_counts %>% 
  arrange(desc(total_sequences)) %>% 
  mutate(pct_UMEND = UMEND / total_sequences,
         Percent_duplicates = 1 - (UMND / UM), 
         not_uniquely_mapped = total_sequences - UM,
         duplicate_reads = UM - UMND,
         non_exonic_reads = UMND - UMEND,
         sum_slices = UMEND + non_exonic_reads + duplicate_reads + not_uniquely_mapped)

```

## Convert to long form

```{r}

plotInfo <- read_counts_with_slices %>% 
			select(sample_id,
			       not_uniquely_mapped,
			       duplicate_reads,
			       non_exonic_reads,
			       UMEND,
			       Percent_duplicates,
			       pct_UMEND,
			       total_sequences) %>%
			gather(read_type, read_count, not_uniquely_mapped, duplicate_reads, non_exonic_reads, UMEND) %>%
			mutate(read_type = intermediate_names_for_read_types[match(read_type, raw_names_for_read_types)],
			       read_type = factor(read_type,
			                          levels = intermediate_names_for_read_types),
			       sample_id = factor(sample_id, levels = orderByTotalReads),
			       read_countM = read_count / 1e6,
			       percent_of_total = read_count / total_sequences)

```

# Make 55 sample subset

```{r}
                 
# smaller dataset to show individual samples
read_counts_from_reasonably_sized_samples <- read_counts %>%
  filter(total_sequences < 150E6) # the most relevant set, and limiting makes plots intelligible


selected_parent_samples <- tibble(THid = parent_sample_data$`Treehouse ID`)

# selected_parent_samples$plot_id = paste0("S", 1:nrow(selected_parent_samples))
selected_parent_samples$plot_id = "*"

n_random_samples_needed <- nrow(read_counts_from_reasonably_sized_samples) - nrow(selected_parent_samples)
set.seed(2)
random_samples <- read_counts_from_reasonably_sized_samples[sample(1:n_random_samples_needed,n_samples_to_plot),]$sample_id
samples_for_small_dataset <- c(selected_parent_samples$THid, random_samples)
# special_designation=c("other", "From_Fig1")
# special_designation <- c("", "*")
small_plot_data <- subset(plotInfo, sample_id %in% samples_for_small_dataset) %>%
  mutate(highlight_text = selected_parent_samples$plot_id[match(sample_id, selected_parent_samples$THid)])
# small_plot_data$highlight_text = factor(special_designation[1 + small_plot_data$sample_id %in% example_samples_to_force_into_small_dataset],
#                                        levels = rev(special_designation))
selected_small_plot_data <- plotInfo %>%
  dplyr::filter(sample_id %in% selected_parent_samples$THid)

```

# # Define reference ranges based on mean and sd of samples with 20M umend reads

```{r}

read_type_mean_and_sd_20U <- plotInfo %>% 
  group_by(sample_id) %>%
  dplyr::filter(read_countM[read_type == "UMEND"] > 20) %>% 
  group_by(read_type) %>% 
  summarize(mean_pct_of_total = mean(percent_of_total), 
            one_sd = sd(percent_of_total),
            mean_minus_one_sd = mean_pct_of_total - one_sd,
            mean_plus_one_sd = mean_pct_of_total + one_sd,
            mean_minus_two_sd = mean_pct_of_total - one_sd * 2,
            mean_plus_two_sd = mean_pct_of_total + one_sd * 2) %>% 
  gather(stat, value, -one_sd, -read_type) %>%
  mutate(stat2 = gsub("^.*two_sd", "two_sd", gsub("^.*one_sd", "one_sd", stat)))

read_types_with_a_lower_bound <- c("UMEND", "Duplicate reads")

read_type_pct_thresholds_raw <- read_type_mean_and_sd_20U %>% 
  dplyr::filter(grepl("two_sd", stat)) %>%
  mutate(value2 = ifelse(grepl("minus_two_sd", stat),
                         ifelse(read_type %in% read_types_with_a_lower_bound,
                                value,
                                0),
                         value))
# 
# read_type_pct_thresholds <- read_type_pct_thresholds_raw %>% 
#  select(read_type, stat, value2) %>%
#   spread(stat, value2) %>%
#   rename(min_pct = mean_minus_two_sd, max_pct = mean_plus_two_sd)

read_type_pct_thresholds <- read_type_pct_thresholds_raw %>% 
 select(read_type, stat, value) %>%
  spread(stat, value) %>%
  rename(min_pct = mean_minus_two_sd, max_pct = mean_plus_two_sd)



```


# Identify samples within reference range

```{r}

plot_info_with_limits <- left_join(plotInfo, read_type_pct_thresholds, by = "read_type") %>%
  mutate(within_limits = ifelse(percent_of_total > min_pct,
                                ifelse(percent_of_total<max_pct,
                                       "within_limits",
                                       "above_limit"),
                                "below_limit"))

samples_within_read_type_limits <- plot_info_with_limits %>% 
  group_by(sample_id) %>%
  filter(sum(grepl("within_limits", within_limits)) == 4) %>%
  pull(sample_id) %>% unique


length(unique(plot_info_with_limits$sample_id))
length(samples_within_read_type_limits)

length(samples_within_read_type_limits)/length(unique(plot_info_with_limits$sample_id))
```

# Annotate small samples data to indicate those not WNL
```{r}

small_plot_data$not_WNL <- small_plot_data$sample_id  %in% samples_within_read_type_limits

threshold_codes=c("NUM"="Not uniquely mapped", "D"="Duplicate reads", "NE" = "Non exonic reads", "U" = "UMEND")

threshold_failure_notes <- subset(plot_info_with_limits, sample_id  %in% subset(small_plot_data, ! not_WNL)$sample_id) %>% 
  filter(within_limits != "within_limits") %>%
  mutate(threshold_code = names(threshold_codes)[match(read_type, threshold_codes)]) %>%
  group_by(sample_id, total_sequences) %>%
  summarize(failed_thresholds = paste(threshold_code, collapse=", "))

# to review visually
subset(plot_info_with_limits, sample_id  %in% subset(small_plot_data, ! not_WNL)$sample_id) %>% 
  filter(within_limits != "within_limits") %>%
  select(sample_id, read_type, within_limits) %>%
  spread(read_type, within_limits) 

# of the samples i've looked at, a below threshold UMEND value is always accompanied by a value that exceeds the threshold in another category


```



# Scale UMEND to 20M

```{r}

# includes all samples

min_UMEND = 20 * 1e6

adjusted_read_counts <- plotInfo %>% 
  group_by(sample_id) %>% 
  mutate(adjustment_factor = 1 / (read_count[read_type == "UMEND"]/min_UMEND),
         adjust_read_count = adjustment_factor * read_count,
         total_adjusted_read_count = sum(adjust_read_count))

```

# Calculate and record stats for scaling UMEND to all depth thresholds
```{r}

# includes only samples within the reference ranges

proposed_thresholds = c(1,seq(4,48, by = 4))
percentiles_to_query = 0.95 # c(50, 75, 95, 100)/100

#library(purrr)

# remove failed samples
plot_info_within_limits <- plotInfo %>%
  filter(sample_id %in% samples_within_read_type_limits)

proposed_threshold=1
percentile_to_query=0.95

rescale_and_get_percentile <- function(proposed_threshold, percentile_to_query) {
  plot_info_within_limits %>% 
    filter(read_type == "UMEND") %>%
    mutate(adjustment_factor = 1 / (read_count / (proposed_threshold * 1e6)),
           adjusted_total = adjustment_factor * total_sequences) %>%
    pull(adjusted_total) %>%
    quantile(percentile_to_query)
}

predictions <- NULL
for (proposed_threshold in proposed_thresholds) {
  for (percentile_to_query in percentiles_to_query){
    predictions <- bind_rows(predictions,
                             tibble(proposed_threshold,
                                    percentile_to_query,
                                    predicted_total_reads = rescale_and_get_percentile(proposed_threshold,
                                                                                       percentile_to_query)))
  }
}


write_tsv(predictions %>% filter(percentile_to_query==0.95),
          "../data/predicted_required_total_reads_within_limits.tsv")



```

# Figure 2

### 2A
```{r}
plot_title <- "Total number of reads in a sample does not predict the number of UMEND reads"
figure_name <- "read_count_and_fractions_figures_v2"


# note negative values in small_plot_data$read_countM, e.g. -0.4817530
p1 <- ggplot(small_plot_data) +
  geom_bar(data = small_plot_data,
           aes(x = sample_id,
               y = read_countM,
               fill = read_type),
           size = 1,
           stat = "identity") +
  scale_fill_manual("Read type", values = category_colors, labels=better_names_for_read_types) +
  scale_color_grey(end = 0.6) +
  # facet_grid(.~highlight_sample, scales="free_x", space="free") + 
  xlab(paste0("Samples, n=", length(samples_for_small_dataset))) +
  ylab("Read count (millions)") +
  ylim(0, 175) +
  geom_text(data=subset(small_plot_data, read_type == "UMEND"),
                               aes(x = sample_id,
                y = 1 + total_sequences / 1e6,
                label = highlight_text),
            color = "black",
            size = 5) +
  geom_text(data=threshold_failure_notes,
            aes(x = sample_id,
                y = 1 + total_sequences / 1e6,
                label = failed_thresholds),
            angle=90, hjust=0,
            color = "darkred",
            size=3) +
  # labs(caption= paste(figure_name, Sys.time())) + 
    theme(axis.ticks.x = element_blank(), 
          axis.text.x = element_blank(),
           legend.position = c(.4,.86)
        #   plot.caption = element_text(hjust = -1)
    ) 
  #geom_text(data=selected_small_plot_data, aes(x=sample_id, y=total_sequences), label="*")

p1


```


### 2B
```{r}

plotInfo_20U <- plotInfo %>% 
  group_by(sample_id) %>%
  dplyr::filter(read_countM[read_type == "UMEND"] > 20)

read_type_names <- better_names_for_read_types
names(read_type_names)= names(category_colors)

read_type_labeller <- function(variable,value){
  return(read_type_names[value])
}


p20 <-  ggplot(plotInfo_20U) + 
  geom_histogram(aes(x = percent_of_total, fill = read_type)) + 
  geom_vline(data = read_type_mean_and_sd_20U, 
             aes(xintercept = value,
                 color = read_type,
                 linetype = stat2)) +
  scale_fill_manual(values = category_colors) +
  scale_color_manual(values = category_colors) +
  facet_wrap(~read_type, nrow=1, labeller=read_type_labeller) 

p20b <- p20 + 
  theme(legend.position = "none") +
  ylab("Samples") + 
  xlab(paste0("Fraction of total reads, n=", length(unique(plotInfo_20U$sample_id)))) 

padding_from_line <- -0.04

read_type_pct_thresholds_for_fig2 <- read_type_pct_thresholds %>%
  gather(threshold, value, -read_type) %>%
  mutate(
    x = ifelse(threshold=="max_pct", value + padding_from_line, value - padding_from_line),
#    y = ifelse(read_type %in% c("Duplicate reads", "UMEND"), 200, 50),
    y = 200,
    hjust = ifelse(threshold=="max_pct", "left", "right"),
    label = ifelse(threshold=="max_pct", 
                   paste0("mean\n+ 2sd\n", round(value, 3)),
                   paste0("mean\n- 2sd\n", round(value, 3))
                   )
    )

p20c <- 
  p20b + 
  geom_label(data = read_type_pct_thresholds_for_fig2, 
            aes(x = x, y = y, label = round(value, 3), hjust = hjust)
  ) + 
  scale_x_continuous(breaks = seq(0, 1, by=0.25)) +
  expand_limits(x = c(-0.25, 0.9))

p20c
```




### Combine figures 2A and 2B
```{r}

#p12 <- plot_grid(p1, p20c, labels = c("A", "B"), rel_widths = c(2.5, 1.5) , axis = "x", align = "h")

p12 <- plot_grid(p1, p20c, labels = c("A", "B"), rel_heights = c(4, 1.5), nrow=2 , axis = "y", align = "v")


#title <- ggdraw() +
 # draw_label(plot_title, fontface = 'bold')

fig_file_base <- paste(figure_name, Sys.time())

#plot_output <- plot_grid(title, p12, ncol = 1, rel_heights = c(0.1, 1)) + 
plot_output <-   p12 +
  draw_label(fig_file_base,
             x = 0,
             y = 0,
             vjust = 0,
             hjust = 0,
             size = 10) #, fontface = 'bold')

plot_output
# ggsave(paste0("../figures_and_tables/", gsub(" ", "_", gsub(":", "-", fig_file_base)), ".png"), plot_output, height = 5, width = 12.5)
ggsave(paste0("../figures_and_tables/", gsub(" ", "_", gsub(":", "-", fig_file_base)), ".png"), plot_output, height = 7, width = 11)
```

# Text related to figure 2
```{r}

number_of_samples_analyzed <-  nrow(complete_counts_raw)


project_ids <- gsub("TARGET-40.*$", "TARGET-40", gsub("_.*", "", complete_counts_raw$sample_id)) 

number_of_projects <- length(unique(project_ids))

print(paste("The total number of samples analyzed was",number_of_samples_analyzed, "from", number_of_projects, "projects"))


number_of_samples_with_more_than_20M_UMEND_reads <- nrow(subset(complete_counts_raw, UMEND> 20E6))
print(paste( number_of_samples_with_more_than_20M_UMEND_reads, "samples have more than 20M UMEND reads"))

number_of_samples_within_reference_ranges <- length(samples_within_read_type_limits)
print(paste(number_of_samples_within_reference_ranges,"samples that had values within the reference range"))

median_num_of_UMEND_reads_in_samples_within_limits <- plot_info_with_limits %>% 
  filter(within_limits == "within_limits",
         read_type == "UMEND") %>%
  pull(read_countM) %>%
  median
  
print(paste("the median number of UMEND reads in these samples was", round(median_num_of_UMEND_reads_in_samples_within_limits, 1), "million"))


# nrow(subset(complete_counts_raw, UMEND> 20E6 & sample_id %in% samples_within_read_type_limits)) # 710
# 
# total_lost_to_10_UMEND_threshold <-  nrow(subset(complete_counts_raw, UMEND< 10E6)) # 710
# 
# pct_lost_to_10_UMEND_threshold <-  total_lost_to_10_UMEND_threshold/number_of_samples_analyzed
# 
# total_lost_to_ref_range <- number_of_samples_analyzed - number_of_samples_within_reference_ranges
# 
# pct_lost_to_ref_range <- total_lost_to_ref_range/number_of_samples_analyzed

```


# Figure 5 - 

```{r}

figure_name <- "predicted_required_total_reads"
fig_file_base <- paste(figure_name, Sys.time())

this_plot_data <- subset(adjusted_read_counts,
                   sample_id %in% samples_for_small_dataset &
                   sample_id %in% samples_within_read_type_limits)

this_plot_data_for_labels <- this_plot_data %>%
  filter(sample_id %in% selected_parent_samples$THid,
         read_type == "UMEND") 

p <- ggplot(this_plot_data) + 
  geom_bar(aes(x = sample_id, 
               y = adjust_read_count / 1e6, 
               fill = read_type), size = 1, stat = "identity") +
  geom_hline(yintercept = 70, linetype = "dashed") +
  scale_fill_manual("Read type", values = category_colors, labels=better_names_for_read_types) +
  scale_color_grey(end = 0.6) + 
  xlab("Sample") +
  theme(axis.ticks.x = element_blank(), 
        axis.text.x = element_blank()) +
  ylab("Scaled read count (million)") + 
  xlab(paste0("Samples, n=", length(unique(this_plot_data$sample_id)))) +
  geom_text(data=this_plot_data_for_labels,
                               aes(x = sample_id,
                y = 1 + total_adjusted_read_count / 1e6,
                label = "*"),
            color = "black",
            size = 5) 
###

plot_output <- plot_grid(p) +
  draw_label(paste(figure_name, Sys.time()),
             x = 0, 
             y = 0,             
             vjust = 0, 
             hjust = 0, 
             size = 10) 

plot_output

ggsave(paste0("../figures_and_tables/", gsub(" ", "_", gsub(":", "-", fig_file_base)), ".png"), plot_output, height = 5, width = 8)
```

This file was saved with the timestamp `r fig_file_base`

# Text around fig 5
```{r}


print(paste(number_of_samples_within_reference_ranges,"samples that had values within the reference range"))

adjusted_UMEND_wnl <- 
adjusted_read_counts %>%
filter(sample_id %in% samples_within_read_type_limits,
       read_type == "UMEND"
       ) 


# t5 is the  nsamples in the reference range require fewer than 70M total reads to contain 20M UMEND reads

t5 <- adjusted_UMEND_wnl %>% filter(total_adjusted_read_count < 70E6) %>% nrow
t5/nrow(adjusted_UMEND_wnl)

print(paste("All samples in this figure and", round(100*t5/nrow(adjusted_UMEND_wnl)), "% of the", number_of_samples_within_reference_ranges, "samples in the reference range require fewer than 70M total reads to contain 20M UMEND reads (dashed horizontal line)."))


```

# QC


## Average duplicate rate per project

```{r}

read_counts_with_slices %>% 
  mutate(project=gsub("[-_].*$", "", sample_id)) %>%
  group_by(project) %>%
  summarize(
    mean_pct_dupe=mean(Percent_duplicates),
    n=n()
    ) %>%
  arrange((mean_pct_dupe))

read_counts_with_slices %>% 
  filter(UMEND>20E6) %>%
  mutate(project=gsub("[-_].*$", "", sample_id)) %>%
  group_by(project) %>%
  summarize(
    mean_pct_dupe=mean(Percent_duplicates),
    n=n()
    ) %>%
  arrange((mean_pct_dupe))



```
