---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
#  pdf_document: default
 # fig_caption: yes

---


```{r parameters}
min_correlation=0.875

neighbor_filter="same_disease"

f_make_supporting_plots=FALSE


```




```{r test_defs}
this_truth_definition="found_in_3_of_4_biggest_subsets"

truth_definitions=c("found_in_at_least_half_the_subsets", 
                    "found_in_the_deepest_subset",
                    "found_in_3_of_4_biggest_subsets"
)
```

In the analysis in this document, truth is defined as "`r gsub("_", " ", this_truth_definition)`"

Specifically, a sample is considered a true "first degree most correlated sample" 
to a parent sample
if it is `r gsub("_", " ", this_truth_definition)` made from that parent.

The available definitions are:

`r paste(truth_definitions, sep="\n")`

```{r setup, include=FALSE}

#knitr::opts_knit$set(root.dir = "~/Documents/Dropbox/ucsc/projects/gitCode/calculations for umend_qc_publication/")



library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(forcats)
library(janitor)
library(grid)
	library(gtable)
	library(gridExtra)

```



                   
## Load data


```{r }


this_bin_width=4

#first_degree_results_with_self_correlations <- 
first_degree_results_with_corrs_gt_875 <- 
  read_tsv("../data/correlated_samples.txt",
  col_types=cols()
  ) %>% 
  rename(neighbor_sample = first_degree_neighbor_sample) %>%
  mutate(
    sample_is_neighbor_in_subset = TRUE, 
    parent_sample=gsub("_est.*$", "", focus_sample),
    target_UMEND_count=as.numeric(gsub("^.*_est([0-9]*)M_.*$", "\\1", focus_sample))
  ) %>%
  group_by(parent_sample, neighbor_sample)

evenly_selected_subsamples <- read_tsv(
  paste0("../data/evenly_selected_subsamples_with_binwidth_", this_bin_width, ".tsv"),
  col_types=cols()
  )

parent_sample_metadata<- read_tsv("../data/parent_sample_data_source.tsv",
  col_types=cols()) %>%
# parent_sample_metadata$`technical_sample_id-qc_only`
rename(treehouse_id=`Treehouse ID`)



disease_info_for_compendium<-read_tsv("../data/TreehousePEDv9_clinical_metadata.2019-03-15.tsv", col_types=cols()) %>% rename(THid=th_sampleid)
# disease_info_for_compendium<-read_tsv("../data/sample_data_for_qc_paper_downloaded_via_API_2019-01-04_07.01.47PM.txt",
#  col_types=cols())

disease_info_for_compendium



```



## Definitions

```{r}
first_degree_neighbor_result_codes=tibble(
	sample_is_true_positive_neighbor_in_parent_sample=c(T,F,T,F) , 
	sample_is_neighbor_in_subset=c(T,T,F,F), 
	hit_type=c("True positive", "False positive", "False negative", "True negative"))


plot_colors=c("False positive"="red", "True negative"="pink", 
              "True positive"="lightblue", "False negative"="blue" )

library(RColorBrewer)
source_palette = brewer.pal(12, "Paired")

cb_dark_orange=source_palette[8]
cb_light_orange=source_palette[7]

cb_dark_green=source_palette[4]
cb_light_green=source_palette[3]

cb_dark_purple=source_palette[10]
cb_light_purple=source_palette[9]

cb_dark_blue=source_palette[2]
cb_light_blue=source_palette[1]

plot_colors=c("False positive"=cb_light_orange, "True negative"=cb_dark_orange, 
              "True positive"=cb_dark_purple, "False negative"=cb_light_purple )

accuracy_line_colors<-c( "sensitivity"=as.character(plot_colors["True positive"]),
                         "specificity"=as.character(plot_colors["True negative"]))

fill_colors=c("TRUE"="black", "FALSE"="white")

```

## Add disease & bin info and exclude self-correlations
```{r}

cohort_sizes_in_compendium <- disease_info_for_compendium %>% 
  group_by (disease) %>%
  summarize(matched_disease_cohort_size = n()) 

first_degree_results_plus <- left_join(
  first_degree_results_with_corrs_gt_875,
  parent_sample_metadata,
  by=c("parent_sample"="Eval_ID")
) %>% 
	left_join(
	  evenly_selected_subsamples %>% select(sample_id, median_of_depth_bin) %>% distinct, 
    by = c("focus_sample"="sample_id")
	) 


first_degree_results <- first_degree_results_plus %>%
  dplyr::filter(neighbor_sample!=treehouse_id) %>%
left_join(disease_info_for_compendium %>% select (THid, disease),
          by=c("neighbor_sample"="THid")) %>%
  select("neighbor_disease"="disease", diagnosis="Disease", everything()) %>%
  left_join(cohort_sizes_in_compendium, 
            by=c("diagnosis"="disease")) %>%
#  select(everything(), focus_disease_cohort_size_in_compendium=matched_disease_cohort_size)
   select(-matched_disease_cohort_size)


```


## Plan X axis values and labels
```{r}

outlier_results_x_breaks=seq(
        0, 
        max(first_degree_results$median_of_depth_bin, na.rm=TRUE),
        by=this_bin_width
      )
outlier_results_x_labels=gsub("^0$", 1, outlier_results_x_breaks)


```



## Apply filters to which neighbors we're analyzing 
```{r}

# if (neighbor_filter=="same_disease"){
  filter_description="Limit analysis to same disease neighbors"
  first_degree_results_filtered<- first_degree_results %>%
    dplyr::filter(diagnosis==neighbor_disease)
# }

number_of_parent_samples=length(unique(first_degree_results_filtered$parent_sample))


```


## Identify true neighbors &  annotate calls

```{r }
annotated_fdn_results_raw <- lapply(unique(first_degree_results_filtered $parent_sample), 
                                    function(this_parent_sample){
	# this_parent_sample =unique(first_degree_results_filtered $parent_sample)[1]
  # this_parent_sample="TH_Eval_019"
	fdn_found_in_selected_subsamples=subset(first_degree_results_filtered,
		 parent_sample == this_parent_sample & 
		 focus_sample %in% evenly_selected_subsamples$sample_id) 
	###
	subsamples <- unique(fdn_found_in_selected_subsamples$focus_sample)
	samples_that_are_fdn_in_any_subset <- unique(fdn_found_in_selected_subsamples$neighbor_sample)
	# add data to show a negative result for genes not found in the outlier list
	generic_subsample_info <- fdn_found_in_selected_subsamples %>% 
	  ungroup %>%
	  select(
	    parent_sample, focus_sample, target_UMEND_count, treehouse_id, diagnosis, publication_id, median_of_depth_bin
	  ) %>% distinct
	
  evenly_selected_subsamples_this_parent	 <- subset(evenly_selected_subsamples, parent_id == this_parent_sample) 
	
	empty_results <- crossing(evenly_selected_subsamples_this_parent, neighbor_sample = samples_that_are_fdn_in_any_subset)
	  
#	  evenly_selected_subsamples_this_parent %>%
#	  mutate(neighbor_sample = rep(samples_that_are_fdn_in_any_subset, each = nrow(evenly_selected_subsamples_this_parent)))
	
	# empty_results <- tibble(parent_sample = this_parent_sample,
	#                         focus_sample = rep(subsamples,
	#                                         each = length(samples_that_are_fdn_in_any_subset)),
	#                         neighbor_sample = rep(samples_that_are_fdn_in_any_subset,
	#                                    length(subsamples)))
	# 
	# 
	
	
	# subset(evenly_selected_subsamples, parent_id == this_parent_sample)
	
	fdn_results <- right_join(fdn_found_in_selected_subsamples %>% select(-target_UMEND_count, -median_of_depth_bin,  -publication_id),
	                          empty_results,
	                          by = c("parent_sample"="parent_id", "focus_sample"="sample_id", "neighbor_sample")) %>%
	  mutate(sample_is_neighbor_in_subset = ifelse(is.na(sample_is_neighbor_in_subset), FALSE, TRUE))
	
	# define true positives

	if(this_truth_definition=="found_in_3_of_4_biggest_subsets"){
	  	
	# define true positives
	biggest_4_subsets=sort(unique(fdn_found_in_selected_subsamples$target_UMEND_count), decreasing=TRUE)[1:4]
	minimum_hits_for_true_neighbors=3
  
    neighbor_frequency<-fdn_found_in_selected_subsamples %>% 
    dplyr::filter(
      sample_is_neighbor_in_subset,
      target_UMEND_count %in% biggest_4_subsets) %>% 
    group_by(neighbor_sample) %>% 
    summarize(n_subsamples_found=length(unique(focus_sample)))
  
  true_neighbors_in_parent_sample <- neighbor_frequency %>% 
    dplyr::filter(n_subsamples_found>=minimum_hits_for_true_neighbors) %>% 
    pull(neighbor_sample)
}

	  if(this_truth_definition=="found_in_the_deepest_subset"){
  	  	
      true_neighbors_in_parent_sample <- subset(first_degree_results_filtered,
  		 parent_sample == this_parent_sample & 
  		 target_UMEND_count==50) %>%
      dplyr::filter(sample_is_neighbor_in_subset) %>% 
      pull(neighbor_sample)
      
      
	  }
	 if(this_truth_definition=="found_in_at_least_half_the_subsets"){
        minimum_hits_for_true_neighbors= length(unique(fdn_found_in_selected_subsamples$focus_sample))/2
  
      neighbor_frequency<-fdn_found_in_selected_subsamples %>% 
      dplyr::filter(sample_is_neighbor_in_subset) %>% 
      group_by(neighbor_sample) %>% 
      summarize(n_subsamples_found_in=length(unique(focus_sample)))
    
      true_neighbors_in_parent_sample <- neighbor_frequency %>% 
      dplyr::filter(n_subsamples_found_in>=minimum_hits_for_true_neighbors) %>% 
      pull(neighbor_sample)
	 }
	
	# except when True=present@50, exclude can2didate neighbors that are never positive in this data set
  if(this_truth_definition!="found_in_the_deepest_subset"){
	  fdn_found_in_selected_subsamples <- fdn_found_in_selected_subsamples %>% 
	  group_by(neighbor_sample) %>%
	  dplyr::filter(sum(sample_is_neighbor_in_subset)>0) 
}
	
	
	this_annotated_fdn_results <-fdn_results %>% 
		group_by(neighbor_sample) %>% 
		mutate(
			neighbor_appearances=sum(as.numeric(sample_is_neighbor_in_subset)),
			first_appears=min(median_of_depth_bin [sample_is_neighbor_in_subset]),
			sample_is_true_positive_neighbor_in_parent_sample= 
			  neighbor_sample %in% 
			    true_neighbors_in_parent_sample,
#			max_corr=max(correlation),
			focus_and_neighbor_have_same_disease=diagnosis==neighbor_disease
		)
	return(this_annotated_fdn_results)
}) %>% bind_rows

table(annotated_fdn_results_raw$sample_is_true_positive_neighbor_in_parent_sample)
table(annotated_fdn_results_raw$sample_is_neighbor_in_subset)



annotated_fdn_results <- annotated_fdn_results_raw %>%
# 		left_join(evenly_selected_subsamples %>% select(sample_id, median_of_depth_bin),  # moved earlier
#		          by = c("focus_sample"="sample_id")
#		) %>%
		left_join(first_degree_neighbor_result_codes, by = 
		            c("sample_is_neighbor_in_subset", "sample_is_true_positive_neighbor_in_parent_sample")
		)


```

## prepare data for plotting
```{r}

# write_tsv(annotated_fdn_results, paste0("data/annotated_1deg_neighbor_results", "-", this_truth_definition, ".txt"))


ordered_plot_input<-  annotated_fdn_results %>% 
  #arrange(sample_is_true_positive_neighbor_in_parent_sample, neighbor_appearances, rev(first_appears)) %>%
  arrange(sample_is_true_positive_neighbor_in_parent_sample, neighbor_appearances) %>%
  ungroup %>%
  mutate(
  	unique_focus_neighbor_id =	paste(parent_sample, neighbor_sample, sep="_"),
 	  ordered_neighbor=factor(unique_focus_neighbor_id, levels=rev(unique(unique_focus_neighbor_id)))
  ) %>% group_by(parent_sample) %>%
  mutate(parent_sample_label=paste0(parent_sample, " (n=", length(unique(unique_focus_neighbor_id)), ")"))


```


# Make multi-sample plot identified by final correlation
```{r}
# 
# 
# 
# 
# 
# these_breaks=c(0,0.88, 0.885, 0.89, 0.895, 0.9, 1)
# these_labels=c("<88%", "<88.5%", "<89%", "<89.5%",  "<90%", ">90%")
# #                       # labels=c("below 88%", "88-88.5%", "88.5-89%", "89-89.5%",  "89.5-90%", "above 90%"),
# 
# ordered_plot_input_corr_code<-ordered_plot_input %>%
#   ungroup %>%
#   mutate(
#     max_corr_group=cut(max_corr, 
#                        breaks=these_breaks,
#                        labels=these_labels
#                        ),
#     current_corr_group=cut(correlation, 
#                        breaks=these_breaks,
#                        labels=these_labels
#                        )
#   )
# 
# # each tile is colored by current correlation
# if(f_make_supporting_plots){
#   ggplot(ordered_plot_input_corr_code, aes(x = median_of_depth_bin, 
#                                 y = ordered_neighbor,
#                                 fill=current_corr_group,
#                                 alpha=sample_is_neighbor_in_subset)) +
#   geom_tile(stat = "identity", size=1) +
#   ylab("candidate neighbor (correlation to focus sample)") +
#   scale_fill_brewer(palette="Set1") + 
#   ggtitle("Molecular similarity analysis results", "grouped by final correlation at max depth") +
#     #plot_title, paste("parent disease is", subset(parent_sample_metadata, sample_id==this_parent_sample)$diagnosis, "\nTruth is defined as", this_truth_definition) ) + # theme(axis.text.y  = element_blank()) +
#   scale_x_continuous("UMEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) +
#     facet_grid(max_corr_group~.,scales="free_y", space="free") +
#     theme(axis.text.y  = element_blank()) +
#     geom_vline(xintercept=outlier_results_x_breaks,
#       color=grey(0.9),
#       size=0.5
#       )
# 
# 
# # grey all grey
# ggplot(ordered_plot_input_corr_code, aes(x = median_of_depth_bin, 
# #                                y = as.character(round(max_corr,3)), 
#                                 y = ordered_neighbor,
#                                 fill=sample_is_neighbor_in_subset)
# #                                color = "gray"
# ) + 
# #                                alpha=sample_is_neighbor_in_subset)) +
# #                                color = hit_type, fill =focus_and_neighbor_have_same_disease
#   geom_tile(stat = "identity", size=1) +
#   geom_vline(xintercept=outlier_results_x_breaks,
#       color=grey(0.9),
#       size=0.5
#       ) +
#   ylab("candidate neighbor (correlation to focus sample)") +
#   scale_fill_manual(values= c("lightgrey", "darkgrey")) +
# #  ggtitle(plot_title, paste("parent disease is", subset(parent_sample_metadata, sample_id==this_parent_sample)$diagnosis, "\nTruth is defined as", this_truth_definition) ) + # theme(axis.text.y  = element_blank()) +
#   scale_x_continuous("UMEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) +
#   # facet_grid(parent_sample_label~.,scales="free_y", space="free") +
#   #  facet_wrap(~max_corr_group, ncol=1, scales="free_y", strip.position="right") + 
#      facet_grid(max_corr_group~.,scales="free_y", space="free") +
#     theme(axis.text.y  = element_blank())
# }
```

##  What percent of correlated samples were detectable with 1 million UMEND reads?
```{r}
 

# how many (real) correlated samples are there? 59
  subset(ordered_plot_input, median_of_depth_bin==0) %>% pull(unique_focus_neighbor_id) %>% unique %>% length

# 2% of (real) correlated samples were detectable with 1 million UMEND reads
subset(ordered_plot_input, median_of_depth_bin==0 & sample_is_true_positive_neighbor_in_parent_sample) %>% tabyl(first_appears)



```


# Make plots

```{r, fig.height=10, fig.width=9}
# 
# for (this_parent_sample in sort(unique(ordered_plot_input$parent_sample))){
#   # this_parent_sample = unique(ordered_plot_input$parent_sample)[1]
#   plot_title=paste0("first deg neighbors of ", this_parent_sample,  "\neach row is a focus-neighbor combination"  )
#   this_plot_data=subset(ordered_plot_input, parent_sample==this_parent_sample)
#   # holding<-this_plot_data %>% ungroup %>% select(neighbor_sample, ordered_neighbor, max_corr) %>% distinct
#   holding<-this_plot_data %>% ungroup %>% select(neighbor_sample, ordered_neighbor) %>% distinct
#   ybreaks=holding$ordered_neighbor
#   names(ybreaks)=paste0(holding$neighbor_sample, " (", as.character(round(holding$max_corr,3)), ")")
#   
#   p<-ggplot(this_plot_data, aes(x = median_of_depth_bin, 
# #                                y = as.character(round(max_corr,3)), 
#                                 y = ordered_neighbor,
#                                 fill=hit_type)) + 
# #                                color = hit_type, fill =focus_and_neighbor_have_same_disease)) + 
#   geom_tile(stat = "identity", size=1) +
#   ylab("candidate neighbor (correlation to focus sample)") +
#   scale_fill_manual(values= plot_colors) +
#   # scale_fill_manual(values= c("black", "white")) +
#   #scale_fill_manual(gsub("_", " ", "focus_and_neighbor\nhave_same_disease"), values= fill_colors) +
#   ggtitle(plot_title, paste("parent disease is", subset(parent_sample_metadata, sample_id==this_parent_sample)$diagnosis, "\nTruth is defined as", this_truth_definition) ) + # theme(axis.text.y  = element_blank()) +
#     scale_x_continuous("UMEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) +
#   facet_grid(parent_sample_label~.,scales="free_y", space="free") +
#   scale_y_discrete(breaks=ybreaks, labels=names(ybreaks))
#   if (f_make_supporting_plots) print(p)
#   this_plot_data %>% 
#     ungroup %>% 
#     select(neighbor_id=neighbor_sample, max_corr, neighbor_disease, ordered_neighbor) %>% 
#     distinct %>%
#     arrange(desc(ordered_neighbor)) %>%
#     select (-ordered_neighbor) %>%
#     kable( digits=3) %>%
#     print
# }
# 
# #  ggsave(paste0("plots/", plot_title, ".png"))


```




```{r fig.height=10, fig.width=9}

## Make summary plots - one line per neighbor per parent
# 
# if(f_make_supporting_plots){
# plot2<-ggplot(ordered_plot_input, aes(x = median_of_depth_bin, y = ordered_neighbor, fill = hit_type)) + 
#   geom_tile(stat = "identity") +
#   scale_fill_manual(values= plot_colors) +
#   theme(axis.text.y  = element_blank()) +
#     scale_x_continuous("UMEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) 
# 
# plot2 + facet_grid(parent_sample~.,scales="free_y", space="free")
# }
# # ggtitle(plot_title) + 

```



## Calculate fractional accuracy per sample

```{r}
frac_accuracy  <-ordered_plot_input %>% 
    group_by(median_of_depth_bin) %>%
#  group_by(parent_sample_label, focus_sample, median_of_depth_bin) %>%
	summarize(
		fp_frac=sum(hit_type=="False positive")/n(),
		tn_frac=sum(hit_type=="True negative")/n(),
		fn_frac=sum(hit_type=="False negative")/n(),
		tp_frac=sum(hit_type=="True positive")/n())
		
conv_table1<-tibble(fraction_type =c("fp_frac", "tn_frac", "fn_frac", "tp_frac"),
	hit_type=c("False positive", "True negative", "False negative", "True positive"))

	
g_frac_accuracy<-gather(frac_accuracy, fraction_type, fraction_amount,
                          -median_of_depth_bin) %>%
#                        -parent_sample_label, -focus_sample, -median_of_depth_bin) %>%
 left_join(conv_table1) %>%
	mutate(hit_type=factor(hit_type, 
	           levels=c("True negative","False positive","False negative", "True positive")))

head(g_frac_accuracy)


```

## Plot fractional accuracy

```{r, fig.height=10}


  
plot_title="neighbors - all parent samples"  
p4<-ggplot(g_frac_accuracy) + 
	geom_bar(aes(x= median_of_depth_bin, y= fraction_amount/number_of_parent_samples, fill= hit_type), stat="identity") +  
	scale_fill_manual(values= plot_colors) +
      scale_x_continuous("UMEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) +
ggtitle(plot_title, paste("Truth is defined as", this_truth_definition))

# if (f_facet_grid) p4 <- p4 + facet_grid(parent_sample_label~.,scales="free_y", space="free")
if(f_make_supporting_plots) p4

```


# Accuracy rates, e.g. True positives etc.

## Calculations


```{r}


colsToGather=c("median_of_depth_bin", "sensitivity", "specificity", "ppv", "npv") #"tp_count_in_subset", "count_of_positives_in_full_set", "fp_count_in_subset", "count_of_negatives_in_full_set", "all_positives_in_subset", "allNegatives", 

total_samples_in_compendium=11456


accuracy_stats<-NULL
for (this_parent_sample in unique(ordered_plot_input$parent_sample)){
# this_parent_sample="TH_Eval_018"
    this_plot_input=subset(ordered_plot_input, parent_sample==this_parent_sample)
    
    
    
    count_of_positives_in_full_set=length(unique(subset(this_plot_input, sample_is_true_positive_neighbor_in_parent_sample )$neighbor_sample))
    count_of_negatives_in_full_set=total_samples_in_compendium-count_of_positives_in_full_set
    #count_of_negatives_in_full_set=round(0.05*total_samples_in_compendium-count_of_positives_in_full_set)
    
    samples_ever_called_as_neighbor=unique(this_plot_input$neighbor_sample)
    
    count_of_neighbors_not_found = total_samples_in_compendium-length(samples_ever_called_as_neighbor)
    
    ### calculate statistics to compare across samples
    accuracy_counts<-this_plot_input %>% 
      group_by(median_of_depth_bin) %>% 
      summarize(all_positives_in_subset=sum(sample_is_neighbor_in_subset), 
                all_negatives_in_subset=sum(!sample_is_neighbor_in_subset)+ count_of_neighbors_not_found, 
                tp_count_in_subset =sum(hit_type=="True positive"), 
                fp_count_in_subset=sum(hit_type=="False positive"),
                partial_tn_count_in_subset =sum(hit_type=="True negative"), 
                total_tn_count_in_subset= partial_tn_count_in_subset + count_of_neighbors_not_found,
                fn_count_in_subset=sum(hit_type=="False negative")
              )
    
    
    accuracy_stats_this_parent <- accuracy_counts %>% 
      mutate(
        sensitivity= tp_count_in_subset/count_of_positives_in_full_set, 
        specificity= total_tn_count_in_subset/count_of_negatives_in_full_set, 
        ppv= tp_count_in_subset/all_positives_in_subset, 
        npv= total_tn_count_in_subset/all_negatives_in_subset
      )
    
    
    fullStatNames=data.frame(short= c("sensitivity", "specificity", "ppv", "npv"), long=c("sensitivity-detect when present", "specificity-exclude when absent", "ppv-positive predictive value", "npv-negative predictive value"))
    
    g_accuracy_stats_this_parent =gather(accuracy_stats_this_parent[, colsToGather], measurement, value, -median_of_depth_bin) %>%
      mutate(parent_sample=this_parent_sample)
    
   accuracy_stats <- bind_rows(accuracy_stats, g_accuracy_stats_this_parent)
}


```

Note `r length(unique(ordered_plot_input$unique_focus_neighbor_id))` samples were ever called as a neighbor.

## sensitivity and specificity per parent sample


```{r, fig.height=10, fig.width=9}

  accuracy_stats_min<-subset(accuracy_stats,measurement %in% c("sensitivity", "specificity") ) %>%
  group_by(median_of_depth_bin, measurement) %>% 
  mutate(
    depth_bin_min_value=min(value),
    point_label=ifelse(depth_bin_min_value==value, 
                   paste0("x=", median_of_depth_bin, " y=", round(value, 3)),
                   "")
  ) 

proposed_thresholds_raw<- tibble(threshold=c(0.995, 0.95), measurement=c("specificity", "sensitivity"), min_reads=20 )


proposed_thresholds <- lapply(unique(accuracy_stats$parent_sample), function(x) {
  proposed_thresholds_raw %>% mutate(parent_sample=x)
  }) %>% bind_rows
if(f_make_supporting_plots){
ggplot(accuracy_stats_min, aes( y= value, color= measurement, x= median_of_depth_bin)) + geom_point() + geom_smooth(se=FALSE) + ggtitle(paste("sensitivity and specificity per parent sample"))  + facet_wrap(parent_sample~measurement, scale="free_y", ncol=2) + 
  scale_color_manual(values= accuracy_line_colors) 
}

```


# Re-calculate accuracy stats - combined across parent samples
```{r}
colsToGather=c("median_of_depth_bin", "sensitivity", "specificity") #, "ppv", "npv") #"tp_count", "count_of_truly_correlated_samples", "fp_count", "count_of_truly_NON_correlated_samples", "positive_calls", "allNegatives", 

count_of_truly_correlated_samples <- ordered_plot_input %>% 
      dplyr::filter(sample_is_true_positive_neighbor_in_parent_sample) %>%
      select(neighbor_sample, parent_sample) %>% 
      distinct() %>%
      nrow

samples_tested_for_correlation_in_assay <-         
  disease_info_for_compendium %>% 
  dplyr::filter(disease %in% ordered_plot_input$diagnosis) %>%
  nrow


count_of_truly_NON_correlated_samples <- samples_tested_for_correlation_in_assay -
  count_of_truly_correlated_samples

# count_of_truly_NON_correlated_samples <- number_of_parent_samples *
#	total_samples_in_compendium - count_of_truly_correlated_samples
        
samples_ever_called_as_neighbor <- ordered_plot_input %>% 
      select(neighbor_sample, parent_sample) %>% 
      distinct() %>%
      nrow
 
      
samples_NEVER_called_as_neighbor <- samples_tested_for_correlation_in_assay - samples_ever_called_as_neighbor
 
    
    ### calculate statistics to compare across samples
    accuracy_counts<-ordered_plot_input %>% 
      group_by(median_of_depth_bin) %>% 
      summarize(positive_calls=sum(sample_is_neighbor_in_subset), 
                negative_calls=sum(!sample_is_neighbor_in_subset) + samples_NEVER_called_as_neighbor, 
                tp_count =sum(hit_type=="True positive"), 
                fp_count=sum(hit_type=="False positive"),
                tn_count= sum(hit_type=="True negative") + samples_NEVER_called_as_neighbor,
                fn_count=sum(hit_type=="False negative")
              )

    accuracy_stats_raw <- accuracy_counts %>% 
      mutate(
        sensitivity= tp_count/count_of_truly_correlated_samples, 
        specificity= tn_count/count_of_truly_NON_correlated_samples, 
        ppv= tp_count/positive_calls, 
        npv= tn_count/negative_calls
      )
    
    
    fullStatNames=data.frame(short= c("sensitivity", "specificity", "ppv", "npv"), long=c("sensitivity-detect when present", "specificity-exclude when absent", "ppv-positive predictive value", "npv-negative predictive value"))
    
accuracy_stats <- gather(accuracy_stats_raw[, colsToGather], measurement, value, -median_of_depth_bin) 

```


# plots for publication


## Summary plots - one line per neighbor

```{r}

plot_title="Summary neighbor plot - one line per neighbor"  

pCIA<-ggplot(ordered_plot_input, aes(x = median_of_depth_bin, y = ordered_neighbor, fill = hit_type)) + 
  geom_tile(stat = "identity") +
  scale_fill_manual("Call", values= plot_colors) +
#  ggtitle(plot_title) + 
  theme(axis.text.y  = element_blank()) +
  scale_x_continuous("UMEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) +
  xlab("UMEND reads (million)") + 
  ylab("Correlated background sample") +
  theme(
    axis.text.y  = element_blank(),
    axis.ticks.y  = element_blank()
  )  + 
  theme(legend.position="none") + 
  facet_grid(publication_id~.,scales="free_y", space="free")
pCIA

# ggsave(paste0("../figures_and_tables/", plot_title,".", Sys.Date(),  ".png"), plot=pCIA, width=3, height=3)

```

## Plot fractional accuracy

```{r}

  
plot_title="neighbors - all parent samples"  
pCFA<- ggplot(g_frac_accuracy) +  # correlation fractional accuracy
	geom_bar(aes(x= median_of_depth_bin, y= fraction_amount, fill= hit_type), stat="identity") +  
	scale_fill_manual("Call", values= plot_colors) +
  scale_x_continuous("UMEND reads (million)", breaks=outlier_results_x_breaks, labels=outlier_results_x_labels) +
  ylab("Fraction of correlated background samples") +
  theme(
 #   axis.text.y  = element_blank(), # un comment when I've fixed the max = 0.8 issue
 #   axis.ticks.y  = element_blank()
  )  + 
  theme(legend.position="none") 

pCFA

```

# Plot accuracy stats - combined across parent samples
```{r}

  accuracy_stats_combined<-subset(accuracy_stats,measurement %in% c("sensitivity", "specificity") ) %>%
  group_by(median_of_depth_bin, measurement) %>% 
  mutate(
    depth_bin_min_value=min(value),
    point_label=ifelse(depth_bin_min_value==value, 
                   paste0("x=", median_of_depth_bin, " y=", round(value, 3)),
                   "")
  ) 

accuracy_stats_combined_with_no_zeroes<- accuracy_stats_combined %>% 
  ungroup %>%
  mutate(median_of_depth_bin = ifelse(median_of_depth_bin==0, 1, median_of_depth_bin))


pCSS<-ggplot(accuracy_stats_combined_with_no_zeroes, aes( y= value, color= measurement, x= median_of_depth_bin)) + 
  geom_point() +
  geom_smooth(se=TRUE) + 
  scale_color_manual(values= accuracy_line_colors)   + 
  theme(legend.position="none") +
  # ylab("") +
  xlab("UMEND reads (million)" ) +
  ylim(0,1.1) +
  scale_y_continuous("", breaks=seq(0,1,0.2))
 
pCSS # correlation sensitivity and specificity

# ggsave(paste0("figures_and_tables/", plot_title,".", Sys.Date(),  ".png"), plot=pCSS, width=3, height=2)

```


## combine plots

```{r}

combined_figure_name="correlated_samples_figures"



pCFA <- pCFA + ggtitle("B") + theme(plot.title = element_text(hjust = 0))
pCSS <- pCSS + ggtitle("C") + theme(plot.title = element_text(hjust = 0))
pCIA <- pCIA + ggtitle("A") + theme(plot.title = element_text(hjust = 0))
 


	grobCFA<-ggplotGrob(pCFA) # correlation Fractional accuracy
	grobCSS<-ggplotGrob(pCSS) # plot of correlation sensitivity and specificity
	grobCIA<-ggplotGrob(pCIA) # plot of individual correlation accuracy
grobCombo <- arrangeGrob(grobs=list(grobCIA, grobCFA, grobCSS), 
                         layout_matrix=rbind(c(1,2),
                                             c(1,3)),
                         bottom = textGrob(paste(combined_figure_name, Sys.time(), "GMT"),
                                           x=0,
                                           just = "left",
                                           gp = gpar(fontsize = 8))
)
grid.newpage()
grid.draw(grobCombo)


# ggsave(file=paste0("../figures_and_tables/", combined_figure_name, ".png"), grobCombo, width=10, height=8)

ggsave(file=paste0("../figures_and_tables/", combined_figure_name, "_", gsub(" ", "_", gsub(":", "-", Sys.Date())), "_", "UTC", ".png"), grobCombo,  width=10, height=8)

```

	
	
# Table of accuracy stats - combined across parent samples

```{r}

accuracy_stats_combo_for_pub <- accuracy_stats_combined %>% 
  select(median_of_depth_bin,measurement, depth_bin_min_value) %>%
  distinct() %>%
  spread(measurement, depth_bin_min_value) 


accuracy_stats_combo_for_pub %>% 
  kable( digits=4)

write_tsv(accuracy_stats_combo_for_pub, "../data/neighbor_accuracy_stats_combo_for_pub.tsv")

```


