---
title: "`r gsub('.Rmd', ' ', gsub('_', ' ', knitr::current_input()))`"
author: "Holly Beale"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
  # pdf_document: default
  # fig_caption: yes
---


# Set submittable print options

```{r}

f_final_draft <- TRUE
notation_text_alpha <- ifelse (f_final_draft, 0, 1 ) 

```


```{r}

this_truth_definition <- "found_in_3_of_4_biggest_subsets"
truth_definitions <- c("found_in_at_least_half_the_subsets",
                       "found_in_the_deepest_subset",
                       "found_in_3_of_4_biggest_subsets")

f_make_supporting_plots <- FALSE
```

In this document, truth is defined as "`r gsub("_", " ", this_truth_definition)`"

Specifically, a gene is considered a true outlier in a parent sample
if it is `r gsub("_", " ", this_truth_definition)` made from that parent.

The available definitions are:

`r paste(truth_definitions, sep = "\n")`

```{r setup, include = FALSE}

#knitr::opts_knit$set(root.dir = "~/Documents/Dropbox/ucsc/projects/gitCode/calculations for umend_qc_publication/")

library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(forcats)

```

## Load data

```{r}

outlier_results_raw <- read_tsv("../data/outliers.txt") %>% mutate(is_outlier_in_subset = TRUE)

this_bin_width = 4

evenly_selected_subsamples <- read_tsv(paste0("../data/evenly_selected_subsamples_with_binwidth_4.tsv"))

```

## Definitions

```{r}

total_genes_measured <- 58582*0.05

resultsCodes <- tibble(gene_is_true_positive_in_parent_sample = c(T, F, T, F),
                       is_outlier_in_subset = c(T,T,F,F),
                       hit_type = c("True positive", "False positive", "False negative", "True negative"))

source_palette <- brewer.pal(12, "Paired")

cb_dark_red <- source_palette[6]
cb_light_red <- source_palette[5]

cb_dark_blue <- source_palette[2]
cb_light_blue <- source_palette[1]

plot_colors <- c("False positive" = cb_light_red,
                "True negative" = cb_dark_red,
                "True positive" = cb_dark_blue,
                "False negative" = cb_light_blue)

outlier_line_colors <- c("sensitivity" = as.character(plot_colors["True positive"]),
                          "specificity" = as.character(plot_colors["True negative"]))


```

## Match samples with bins

```{r}


annotated_outlier_results <- lapply(unique(outlier_results_raw$parent_sample), function(this_parent_sample) {
	# this_parent_sample <- unique(outlier_results_raw$parent_sample)[1]
	outliers_found_in_selected_subsamples <- subset(outlier_results_raw,
	                                                parent_sample == this_parent_sample &
	                                                  sample_id %in% evenly_selected_subsamples$sample_id)
	subsamples <- unique(outliers_found_in_selected_subsamples$sample_id)
	genes_that_are_outliers_in_any_subset <- unique(outliers_found_in_selected_subsamples$gene)
	# add data to show a negative result for genes not found in the outlier list
	empty_results <- tibble(parent_sample = this_parent_sample,
	                        sample_id = rep(subsamples,
	                                        each = length(genes_that_are_outliers_in_any_subset)),
	                        gene = rep(genes_that_are_outliers_in_any_subset,
	                                   length(subsamples)))
	outlier_results <- left_join(empty_results,
	                             outliers_found_in_selected_subsamples,
	                             by = c("parent_sample", "sample_id", "gene")) %>%
	  mutate(target_MEND_count = as.numeric(gsub("^.*_est", "",
	                                              gsub("M_MEND_seqtk_seed_.*", "", sample_id))),
	         is_outlier_in_subset = ifelse(is.na(is_outlier_in_subset), FALSE, TRUE))
	
	# define true positives

	if (this_truth_definition == "found_in_3_of_4_biggest_subsets") {
  	biggest_4_subsets <- sort(unique(outlier_results$target_MEND_count), decreasing = TRUE)[1:4]
  	minimum_hits_for_true_outlier <- 3
    
    outlier_frequency <- outlier_results %>% 
      dplyr::filter(is_outlier_in_subset,
                    target_MEND_count %in% biggest_4_subsets) %>% 
      group_by(gene) %>% 
      summarize(n_subsamples_found = length(unique(sample_id)))
    
    true_outliers_in_parent_sample <- outlier_frequency %>% 
      dplyr::filter(n_subsamples_found >= minimum_hits_for_true_outlier) %>% 
      pull(gene)
  }
		
  if (this_truth_definition == "found_in_the_deepest_subset") {
    true_outliers_in_parent_sample <- subset(outlier_results_raw,
                                             parent_sample == this_parent_sample &
                                               as.numeric(gsub("^.*_est", "",
                                                               gsub("M_MEND_seqtk_seed_.*",
                                                                    "",
                                                                    sample_id))) == 50) %>%
      dplyr::filter(is_outlier_in_subset) %>%
      pull(gene)
  }

 	if (this_truth_definition == "found_in_at_least_half_the_subsets") {
    minimum_hits_for_true_outlier <- length(unique(outlier_results$sample_id)) / 2

    outlier_frequency <- outliers_found_in_selected_subsamples %>%
      dplyr::filter(is_outlier_in_subset) %>%
      group_by(gene) %>%
      summarize(n_subsamples_found_in = length(unique(sample_id)))

    true_outliers_in_parent_sample <- outlier_frequency %>%
      dplyr::filter(n_subsamples_found_in >= minimum_hits_for_true_outlier) %>%
      pull(gene)
  }
	
	# except when True = present@50, exclude candidate outliers that are never positive in this data set
  if (this_truth_definition != "found_in_the_deepest_subset") {
	  outlier_results <- outlier_results %>% 
  	  group_by(gene) %>%
  	  dplyr::filter(sum(is_outlier_in_subset) > 0) 
  }
	                          
	 this_annotated_outlier_results <- outlier_results %>% group_by(gene) %>% 
		mutate(up_outlier_occurances = sum(as.numeric(is_outlier_in_subset)),
		       first_appears = min(target_MEND_count[is_outlier_in_subset]),
		       gene_is_true_positive_in_parent_sample = gene %in% true_outliers_in_parent_sample) %>%
	   left_join(resultsCodes, by = c("is_outlier_in_subset", "gene_is_true_positive_in_parent_sample"))
	 return(this_annotated_outlier_results)
}) %>% bind_rows

annotated_outlier_results_intermediate <- annotated_outlier_results %>%
  left_join(evenly_selected_subsamples %>% select(-target_MEND_count), 
            by = "sample_id")
annotated_outlier_results$bin_median <- evenly_selected_subsamples$median_of_depth_bin[match(
  annotated_outlier_results$sample_id, evenly_selected_subsamples$sample_id)] # should this line be modifying annotated_outlier_results_intermediate, not annotated_outlier_results?

ordered_plot_input <- annotated_outlier_results_intermediate %>% 
  arrange(gene_is_true_positive_in_parent_sample, up_outlier_occurances, rev(first_appears)) %>%
  ungroup %>%
  mutate(unique_parent_gene_id = paste(parent_sample, gene, sep = "_"),
         ordered_gene = factor(unique_parent_gene_id, levels = rev(unique(unique_parent_gene_id)))) %>%
  group_by(parent_sample) %>%
  mutate(parent_sample_label = paste0(parent_sample, " (n=", length(unique(unique_parent_gene_id)), ")"))

#head(ordered_plot_input)



```


# gather parent outliers
```{r}

parent_metadata <- read_tsv("../data/parent_sample_data_source.tsv")


parent_outliers <- annotated_outlier_results %>% 
  filter(gene_is_true_positive_in_parent_sample) %>% 
  select(Eval_ID=parent_sample, gene) %>%
  distinct %>% 
  left_join(parent_metadata %>% select(Eval_ID, Disease), by=c("Eval_ID")) %>%
  select(Disease, everything()) %>%
  arrange(Disease, Eval_ID, gene)

write_tsv(parent_outliers, "../ancillary/parent_outliers.tsv")

```

# look for literature connecting outliers to genes
```{r}

library("reutils")

control_values = tibble(Disease = "synovial sarcoma", gene = "brca1")



literature_search_inputs <- parent_outliers %>% 
  select(Disease, gene) %>% 
  distinct %>%
  bind_rows(control_values, .) %>%
  mutate(query_text = paste0(Disease," AND expression AND ", gene, "[Gene]"))

# "COI[Gene] AND Gastropoda[ORGN]"

```

```{r}



```

```{r}
reutils.api.key=api_key="4dbd53170c366f5822df4c854696f99f1808"
set_entrez_key(reutils.api.key)
Sys.getenv("ENTREZ_KEY")
```


```{r}
# install.packages("rentrez")
library("rentrez")


# https://ropensci.org/tutorials/rentrez_tutorial/


pubmed_results_list <- lapply(literature_search_inputs$query_text, function(query_t){
  # query_t = "Tanaka TU Stark"
  # query_t = literature_search_inputs$query_text[2]
  # query_t = "breast cancer brca1"
  # query_t = "synovial sarcoma brca1"
  # query_t = "embryonal rhabdomyosarcoma AND DES[Gene]"
  print(query_t)
  this_search <- entrez_search(db="pubmed", 
                               term=query_t,
                               retmax=20)
  Sys.sleep(0.2)
  if (this_search$count>0){
    multi_summs <- entrez_summary(db="pubmed", id=this_search$ids)
    
    t5 <- extract_from_esummary(multi_summs, c("title", "source", "pmcrefcount", "pubdate" )) %>%
      t %>% 
      apply(1, as_tibble) 
    
    t6 <- lapply(t5,  function(x) mutate_all(x, as.character))  %>%
      bind_rows(.id = "PMid" ) %>%
      mutate(query_text = query_t) %>%
      select(query_text, everything())
    
      return(t6)
  } else {
#    return(tibble(PMid = NA,  pubdate= NA, pmcrefcount = NA, title = NA, query_text = query_t))
        return(tibble(query_text = query_t))
  }
}) 

# %>% bind_rows

if (is.null(dim(pubmed_results_list))){
  pubmed_results <- lapply(pubmed_results_list, function(x) mutate_all(x, as.character)) %>% bind_rows
} else {
  pubmed_results <- pubmed_results_list
}  


pubmed_results_no_controls <- pubmed_results %>%
  filter(
    ! query_text == "synovial sarcoma brca1",
    ! grepl(" DES[Gene]", query_text)
    ) %>%
    arrange(query_text, desc(pmcrefcount)) %>%
  rename(time_article_has_been_cited=pmcrefcount)
  
subset(pubmed_results_no_controls,! is.na(pubdate)) %>% View


write_tsv(pubmed_results_no_controls, "../ancillary/all_pubmed_search_results_for_expression_and_outliers_by_disease.txt")  
write_tsv(subset(pubmed_results_no_controls,! is.na(pubdate)), "../ancillary/positive_pubmed_search_results_for_expression_and_outliers_by_disease.txt")  
#        
#        (pubmed_results)
# can't be converted from integer to character
# esummary(this_search)
#  
#  
# r_search$ids

```

```{r}
search_year <- function(year, term){
    query <- paste(term, "AND (", year, "[PDAT])")
    entrez_search(db="pubmed", term=query, retmax=0)$count
}

year <- 2008:2018
papers <- sapply(year, search_year, term="synovial sarcoma", USE.NAMES=FALSE)

plot(year, papers, type='b', main="synovial sarcoma")
```

